Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-05-07 main.Rmd
working  on section 4

# Cyclic calendar algebra {#sec:cyclic-calendar}

In Section \ref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed. In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities <!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated. Since, our method caters to computation of cyclic granularities, we shall refer to it as--> through cyclic calendar algebra operations.

The cyclic calendar algebra consists of broadly two kinds of operations:
(1) **single-to-multiple** and (2) **multiple-to-single** which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively. The hierarchical structure of time creates a natural nested ordering which can produce single-order-up or multiple-order-up granularities. We shall use the notion of a hierarchy table and order to define them.

**Order** of a linear granularity can be comprehended as the level of coarseness associated with a linear granularity. For example, for two linear granularities $G$ and $H$, if $G$ _groups into_ or _finer than_ $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$. In any hierarchy table, linear granularities are arranged from lowest to highest order.

**Hierarchy table** Let $H_n: (G, C, K)$ be a hierarchy table containing $n$ linear granularities. $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$. $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$. We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular. It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.

>>>>

# Hierarchy and ordering {#sec:cyclic-calendar}

<!-- In Section \ref{sec:cyclic-gran}, we discussed how we can define cyclic granularities by extending the framework of linear granularities defined in @Bettini1998-ed. In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities --><!--This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated. Since, our method caters to computation of cyclic granularities, we shall refer to it as--> <!-- through cyclic calendar algebra operations.-->

The cyclic calendar algebra consists of broadly two kinds of operations:
(1) **single-to-multiple** and (2) **multiple-to-single** \todo[inline]{these don't make sense to me, they are just the reverse of each other, so why are both needed. its not  clear from the  next block of text what each operation does} which entails the representation of _multiple-order-up_ cyclic granularities from _single-order-up_ cyclic granularities respectively. The hierarchical structure of time creates a natural nested ordering which can produce single-order-up or multiple-order-up granularities. We shall use the notion of a hierarchy table and order to define them.

\begin{definition}\label{def:order}
The {\bf order} of a linear granularity is the level of coarseness associated with a linear granularity. For example, for two linear granularities $G$ and $H$, if $G$ {\em groups into} or {\em finer than} $H$ or $H$ is composed of higher number of granules of the bottom granularity than $G$, then $G$ is of lower order than $H$. In any hierarchy table, linear granularities are arranged from lowest to highest order.
\end{definition}

\begin{definition}\label{def:hierarchy} 
Let $H_n: (G, C, K)$ be a {\bf hierarchy table} containing $n$ linear granularities. $G_{l}$ and $G_{m}$ represent the linear granularity of order $l$ and $m$ respectively with $l<m$. $K \equiv P(l,m)$ represents the period length of the grouping $(G_{l}, G_{m})$ for a circular granularity and $K \equiv k(l,m)$ represents the operation to obtain $G_{m}$ from $G_{l}$ for a quasi-circular granularity $C_{G(l),G(m)}$, $\forall l,m \in {1, 2, \dots,n}$. We assume that the cyclic granularity $C_{G(l),G(m)}$ is either circular or quasi-circular. It is difficult to position an aperiodic linear granularity in a hierarchy and hence the possibility of an aperiodic cyclic granularity is not considered.
\end{definition}
